Geospatial - Assignment 1 - Moritz Peist, Noemi Lucchi, Simon Vellin

### 2025-01-23 hw1/submission_notebook.Rmd
Paper 1 & 2 done with caching

```{r}
# Dams, Source:https://www.dws.gov.za/DSO/publications.aspx
# Data was only available as .KMZ and not readable by st_read so needed to be converted by:
# https://mygeodata.cloud/converter/kml-to-geojson
dams <- st_read(
  "data/paper1/Registered Dams Oct20241.geojson"
)

@@

# RSA boundaries - downloaded from GDAM and temporarily stored
sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3, path = tempdir()))

# River data - Download elevation data for South Africa
elevation <- elevation_30s(country = "ZAF", path = tempdir())
```

@@

```{r}
# Get Ethiopia districts
eth_districts_sf <- gadm(country = "ETH", level = 3, path = tempdir())
eth_districts_sf <- st_as_sf(eth_districts_sf)

# Get population data and process it for Ethiopia
ethiopia_pop <- population(2020, res = 5, path = tempdir())
ethiopia_pop <- crop(ethiopia_pop, ext(eth_districts_sf))
ethiopia_pop <- mask(ethiopia_pop, vect(eth_districts_sf))

@@

```{r}
# Create the final plot
ggplot() +
  # Add Ethiopia districts with population density fill
  geom_sf(
    data = eth_districts_sf,
    aes(fill = pop_density),
    color = "gray50",
    linewidth = 0.1
  ) +
  # Add power lines in red
  geom_sf(
    data = hv_lines_clean,
    color = "red",
    linewidth = 0.35,
    size = 0.35
  ) +
  # Add major roads in black
  geom_sf(
    data = roads_clipped,
    color = "black",
    linewidth = 0.5,
    size = 0.5
  ) +
  # Add villages as points
  geom_sf(
    data = wb_data_geo,
    color = "black",
    size = 1,
    alpha = 0.6,
    shape = 16 # solid circle
  ) +
  # Customize the fill scale to match paper's blue shading
  scale_fill_gradientn(
    colors = c("white", "lightblue", "royalblue", "navy"),
    trans = "log",
    name = "Population\nDensity\n(people/km²)",
    labels = scales::comma
  ) +
  theme_minimal() +
  labs(
    title = "High Voltage Power Grid Network and Major Roads in Ethiopia",
    subtitle = "Showing transmission lines ≥66kV, population density by district, and villages",
    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank\n
      Source: Based on Fried and Lagakos (2021)"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 8),
    plot.caption = element_text(size = 6)
  )
```

@@

The different color labelling of the surveyed villages is unfortunately not explained in the paper. Therefore, we leave the points black.

>>>>

```{r}
# Set geodata default path
geodata_path("data/paper1")

# Dams, Source:https://www.dws.gov.za/DSO/publications.aspx
# Data was only available as .KMZ and not readable by st_read so needed to be converted by:
# https://mygeodata.cloud/converter/kml-to-geojson
dams <- st_read(
  "data/paper1/Registered Dams Oct20241.geojson"
)

@@

# RSA boundaries - downloaded from GDAM and temporarily stored
sa_districts_sf <- st_as_sf(gadm(country = "ZAF", level = 3))

# River data - Download elevation data for South Africa
elevation <- elevation_30s(country = "ZAF")
```

@@

```{r}
# Set geodata default path
geodata_path("data/paper2")

# Get Ethiopia districts
eth_districts_sf <- gadm(country = "ETH", level = 3)
eth_districts_sf <- st_as_sf(eth_districts_sf)

# Get population data and process it for Ethiopia
ethiopia_pop <- population(2020, res = 5)
ethiopia_pop <- crop(ethiopia_pop, ext(eth_districts_sf))
ethiopia_pop <- mask(ethiopia_pop, vect(eth_districts_sf))

@@

```{r}
# Create the final plot
ggplot() +
  geom_sf(
    data = eth_districts_sf,
    aes(fill = pop_density),
    color = "gray50",
    linewidth = 0.1
  ) +
  scale_fill_gradientn(
    colors = c("white", "lightblue", "royalblue", "navy"),
    trans = "log",
    name = "Population\nDensity\n(people/km²)",
    labels = scales::comma,
    guide = guide_colorbar(order = 1)
  ) +
  geom_sf(
    data = roads_clipped,
    aes(color = "Highway"),
    linewidth = 0.5
  ) +
  geom_sf(
    data = hv_lines_clean,
    aes(color = "HV grid"),
    linewidth = 0.35
  ) +
  geom_sf(
    data = wb_data_geo,
    aes(shape = "Survey village"),
    color = "black",
    size = 1,
    alpha = 0.6
  ) +
  scale_color_manual(
    name = NULL,
    values = c("Highway" = "black", "HV grid" = "red"),
    guide = guide_legend(order = 2)
  ) +
  scale_shape_manual(
    name = NULL,
    values = c("Survey village" = 16),
    guide = guide_legend(order = 3)
  ) +
  guides(
    fill = guide_colorbar(order = 1),
    color = guide_legend(
      order = 2,
      override.aes = list(
        linewidth = 1,
        shape = NA, # Removes the box
        linetype = 1
      )
    ),
    shape = guide_legend(order = 3)
  ) +
  theme_minimal() +
  labs(
    title = "High Voltage Power Grid Network and Major Roads in Ethiopia",
    subtitle = "Showing transmission lines ≥66kV, population density by district, and villages",
    caption = "Data sources: OpenStreetMap, WorldPop 2020, World Bank (2012)\n
      Based on Fried and Lagakos (2021)"
  ) +
  theme(
    legend.position = "left",
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 8),
    plot.caption = element_text(size = 6),
    axis.text = element_blank(),
    axis.title = element_blank()
  )
```

@@

The different color labeling of the surveyed villages is unfortunately not explained in the paper. Therefore, we leave the points black.
